---
title: "R_forecasting_course"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(forecast)

```

# Introduction

## Create time series objects

```{r cars}
# Read the data from Excel into R
mydata <- read_excel("exercise1.xlsx")

# Look at the first few lines of mydata
head(mydata)

# Create a ts object called myts
myts <- ts(mydata[,2:4], start = c(1981, 1), frequency = 4)


```

## Time series plots

You can use the autoplot() function to produce a time plot of the data with or without facets,

You can also use the which.max() and frequency() functions to identify the smallest index of the maximum value and find the number of observations per unit time, respectively. 
Apply the frequency() function to each commodity to get the number of observations per unit time. This would return 52 for weekly data, for example.

```{r pressure, echo=FALSE}

# Plot the data with facetting
autoplot(myts, facets = T)

# Plot the data without facetting
autoplot(myts, facets = F)

# Plot the three series
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)

# Find the outlier in the gold series
goldoutlier <- which.max(gold)

# Look at the seasonal frequencies of gold, woolyrnq and gas
frequency(gold)
frequency(woolyrnq)
frequency(gas)

```

## Seasonal Plots

A seasonal plot is similar to a time plot except that the data are plotted against the individual “seasons” in which the data were observed.
A subseries plot comprises mini time plots for each season. Here, the mean for each season is shown as a blue horizontal line.
One way of splitting a time series is by using window() function, which extracts a subset from the object x observed between the times start and end.

```window(x, start = NULL, end = NULL)```

```{r}

# Load the fpp2 package
library(fpp2)

# Create plots of the a10 data
autoplot(a10)
ggseasonplot(a10)

# Produce a polar coordinate season plot for the a10 data
ggseasonplot(a10, polar = T)

# Restrict the ausbeer data to start in 1992
beer <- window(ausbeer, start=1992)

# Make plots of the beer data
autoplot(beer)
ggsubseriesplot(beer)

```

## Autocorrelation of non-seasonal time series

Another way to look at time series data is to plot each observation against another observation that occurred some time previously by using gglagplot(). For example, you could plot ytyt against yt−1yt−1. 

The correlations associated with the lag plots form what is called the autocorrelation function (ACF). The ggAcf() function produces ACF plots.

```{r}

# Create an autoplot of the oil data
autoplot(oil)

# Create a lag plot of the oil data
gglagplot(oil, k = 1:9)

# Create an ACF plot of the oil data
ggAcf(oil)

```

## Autocorrelation of seasonal and cyclic time series

When data are either seasonal or cyclic, the ACF will peak around the seasonal lags or at the average cycle length.

```{r}
# Plot the annual sunspot numbers
autoplot(sunspot.year)
ggAcf(sunspot.year)

# Save the lag corresponding to maximum autocorrelation
maxlag_sunspot <- 1

# Plot the traffic on the Hyndsight blog
autoplot(hyndsight)
ggAcf(hyndsight)


# Save the lag corresponding to maximum autocorrelation
maxlag_hyndsight <- 7

```

## Stock Prices and White Noise

As you learned in the video, white noise is a term that describes purely random data. You can conduct a Ljung-Box test using the function below to confirm the randomness of a series; a p-value greater than 0.05 suggests that the data are not significantly different from white noise.

There is a well-known result in economics called the "Efficient Market Hypothesis" that states that asset prices reflect all available information. A consequence of this is that the daily changes in stock prices should behave like white noise (ignoring dividends, interest rates and transaction costs). The consequence for forecasters is that the best forecast of the future price is the current price.

You can test this hypothesis by looking at the goog series

```{r}

# Plot the original series
autoplot(goog)

# Plot the differenced series
autoplot(diff(goog))

# ACF of the differenced series
ggAcf(diff(goog))

# Ljung-Box test of the differenced series
Box.test(diff(goog), lag = 10, type = "Ljung")

```

