---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
  word_document: default
---

# Topic

Use twitteR package to load tweets with #cilantro and #coriander - a divisive topic.
Analyse sentiments using __ package

# Load packages etc.

```{r}
library(knitr)

opts_chunk$set(fig.width=6.7, dpi = 300, warning = FALSE, message = FALSE, echo = TRUE)

```

```{r}
library(devtools)
# install.packages("ROAuth")
# install.packages('httr')
# devtools::install_github("geoffjentry/twitteR")
library(twitteR)
library(ROAuth) #for authentication with twitter server
library(magrittr)
library(dplyr)

```

# Create Twitter App

I found all these steps from [rStatistics.net](http://rstatistics.net/extracting-tweets-with-r/)

First, go to [https://apps.twitter.com/](https://apps.twitter.com/)

Set up an app
Click on Keys and Access Tokens tab
record for later tweet uses:

* Consumer Key (API key): mXNImkM2d0DUkM2XA27Xdqb4y 
* Consumer Secret (API secret): LFiGbhAFxnNLq1ycbgosKVTt7CFsX5Cg15gqYnuQx4BXkBKl45 
* Access Token: 2493913478-bqHtbL9LrR3D3fKVe6wWOVfvh6AYzfMEjBX9cfa 
* Access Token Secret: eSZeMMTnPD6DIvuriKPoWC7e2kCXVzx9KoXkpJUgKCHJp 

Authenticate via R:
```{r}
# setup_twitter_oauth(consumerKey, consumerSecret, accessToken, accessTokenSecret)

setup_twitter_oauth('mXNImkM2d0DUkM2XA27Xdqb4y',
                    'LFiGbhAFxnNLq1ycbgosKVTt7CFsX5Cg15gqYnuQx4BXkBKl45',
                    '2493913478-bqHtbL9LrR3D3fKVe6wWOVfvh6AYzfMEjBX9cfa', 
                    'eSZeMMTnPD6DIvuriKPoWC7e2kCXVzx9KoXkpJUgKCHJp')

```

# Retrieve tweets

I want any tweets that contain 'coriander' or 'cilantro'. Cilantro is American and they can be pretty opinionated :p

## General Commands

Before I retrieve mine, some example retrieving commands below:

Tweets from a user or your own account:
```
userTimeline('r_programming',n=10) # tweets from a user

homeTimeline(n=15) # get tweets from home timeline

mentions(n=15) # get your tweets that were retweeted

favs <- favorites("r_programming", n =10) # tweets a user has favorited
```

Tweets with a particular hashtag 

```
tweets <- searchTwitter("rstats", n=25) # top 25 tweets that contain search term

tweetsDF <- twListToDF(tweets) # more info about tweets - converts to nice df
```

## Back to the coriander problem:

Extract tweets and convert to a nice dataframe:

```{r}

tweets <- searchTwitter(c("coriander|cilantro"), 
                        n=500, lang = "en") # top 25 tweets that contain search term
head(tweets)

# extract more info about tweets and convert to a nice df
tweetsDF <- twListToDF(tweets) 

tweetsDF <- tweetsDF %>% select(text, favorited, favoriteCount, created, retweetCount, retweeted, location, language, screenName)

str(tweetsDF)

```

Clean up the text column to remove hashtags, urls, and punctuation

```{r}
text.clean <- function(x) {

  x = gsub('http\\S+\\s*', '', x) ## Remove URLs

  x = gsub('\\b+RT', '', x) ## Remove RT

  x = gsub('#', '', x) ## Remove Hashtags

  x = gsub('@\\S+', '', x) ## Remove Mentions

  x = gsub('[[:cntrl:]]', '', x) ## Remove Controls and special characters

  x = gsub('[,.:;+-]', ' ', x) ## space replaces some Punctuation
  
  x = gsub('[[:punct:]]', '', x) ## Remove Punctuations

  x = gsub("^[[:space:]]*","",x) ## Remove leading whitespaces

  x = gsub("[[:space:]]*$","",x) ## Remove trailing whitespaces

  x = gsub(' +',' ',x) ## Remove extra whitespaces
  
  return(x)

}

#before cleaning
head(tweetsDF$text)
tweetsDF$text <- text.clean(tweetsDF$text)

#after cleaning
head(tweetsDF$text)

```








