---
title: "d3js"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's Do This

```{r cars, echo = F, results = "asis"}
library(jsonlite)
cat(
    paste(
        '<script>
            var data = ',toJSON(cars),';
        </script>'
        , sep = "")
    )
```

<script src="https://d3js.org/d3.v4.min.js"></script>

<div id="plot1"></div>
<div id="plot2"></div>

<script>

// set the dimensions and margins of the graph
var margin = {top: 20, right: 20, bottom: 30, left: 50},
    width = 900 - margin.left - margin.right,
    height = 500 - margin.top - margin.bottom;

// set the ranges
var x = d3.scaleLinear().range([0, width]);
var y = d3.scaleLinear().range([height, 0]);

// append the svg object to the body of the page
// appends a 'group' element to 'svg'
// moves the 'group' element to the top left margin
var svg = d3.select("#plot1").append("svg")
    .attr("width", width + margin.left + margin.right)
    .attr("height", height + margin.top + margin.bottom)
  .append("g")
    .attr("transform",
          "translate(" + margin.left + "," + margin.top + ")");


// format the data
data.forEach(function(d) {
    d.speed = +d.speed;
    d.dist  = +d.dist;
});

// Scale the range of the data
x.domain(d3.extent(data, function(d) { return d.speed; }));
y.domain([0, d3.max(data, function(d) { return d.dist; })]);

// Add the scatterplot
svg.selectAll("dot")
    .data(data)
  .enter().append("circle")
    .attr("r", 5)
    .attr("cx", function(d) { return x(d.speed); })
    .attr("cy", function(d) { return y(d.dist); });

// Add the X Axis
svg.append("g")
    .attr("transform", "translate(0," + height + ")")
    .call(d3.axisBottom(x));

// Add the Y Axis
svg.append("g")
    .call(d3.axisLeft(y));

</script>
</body>


# Separate this shit
```{r}

print('why')
```


<meta charset="utf-8">
<style>

button {
  position: absolute;
  left: 10px;
  top: 10px;
}

</style>
<button onclick="transition()">Update</button>
//<svg width="960" height="500"></svg>

<script>

var n = 20, // number of layers
    m = 200, // number of samples per layer
    k = 10; // number of bumps per layer

var stack = d3.stack().keys(d3.range(n)).offset(d3.stackOffsetWiggle),
    layers0 = stack(d3.transpose(d3.range(n).map(function() { return bumps(m, k); }))),
    layers1 = stack(d3.transpose(d3.range(n).map(function() { return bumps(m, k); }))),
    layers = layers0.concat(layers1);

var svg = d3.select("#plot2").append("svg")
    .attr("width", width + margin.left + margin.right)
    .attr("height", height + margin.top + margin.bottom)

var x = d3.scaleLinear()
    .domain([0, m - 1])
    .range([0, width]);

var y = d3.scaleLinear()
    .domain([d3.min(layers, stackMin), d3.max(layers, stackMax)])
    .range([height, 0]);

var z = d3.interpolateCool;

var area = d3.area()
    .x(function(d, i) { return x(i); })
    .y0(function(d) { return y(d[0]); })
    .y1(function(d) { return y(d[1]); });

svg.selectAll("path")
  .data(layers0)
  .enter().append("path")
    .attr("d", area)
    .attr("fill", function() { return z(Math.random()); });

function stackMax(layer) {
  return d3.max(layer, function(d) { return d[1]; });
}

function stackMin(layer) {
  return d3.min(layer, function(d) { return d[0]; });
}

function transition() {
  var t;
  d3.selectAll("path")
    .data((t = layers1, layers1 = layers0, layers0 = t))
    .transition()
      .duration(2500)
      .attr("d", area);
}

// Inspired by Lee Byron’s test data generator.
function bumps(n, m) {
  var a = [], i;
  for (i = 0; i < n; ++i) a[i] = 0;
  for (i = 0; i < m; ++i) bump(a, n);
  return a;
}

function bump(a, n) {
  var x = 1 / (0.1 + Math.random()),
      y = 2 * Math.random() - 0.5,
      z = 10 / (0.1 + Math.random());
  for (var i = 0; i < n; i++) {
    var w = (i / n - y) * z;
    a[i] += x * Math.exp(-w * w);
  }
}

</script>


